The five axes of the Turtle: symmetry and asymmetry in NGC~6210

### 2020-06-19 ms.tex
Address most of Michael's annotations from a few weeks ago

\begin{abstract}
  We carry out a comprehensive kinematic and morphological study
  of the spectacularly asymmetrical planetary nebula:
  NGC~6210, known as the Turtle Nebula.
  We study the three-dimensional structure and kinematics of the nebula's shells, lobes, knots, and haloes
  by combining radial velocity mapping from multiple long-slit spectra
  with proper motion measurements from multi-epoch imaging.
  We find that the nebula has five separate ejection axes, which we label A~to~E.\@
  Axes~A and B are major axes of the bipolar wind-blown inner shell
  and the lop-sided, elliptical, fainter, but more massive intermediate shell, respectively.
  Axes~C and D (and also A) are bipolar flow axes that form the point symmetric,
  high-ionization outer lobes, all with inclinations close to the plane of the sky.
  Axis~E, which is inclined close to the line of sight,
  is the axis of a loosely collimated outflow of low-ionization knots.
  % A comprehensive kinematic study has been carried out on the planetary nebula (PN) NGC 6210. Multiple long-slits, echelle spectra have been obtained over the face of this nebula mapping its full shell structure, the opposite pairs of extended, twisted arms and the bipolar collimated outflows and bullets. Public {\it HST} imagery has been used to identify kinematic elements with structural components. The long-slit spectroscopic information has been combined into channel maps that greatly facilitate visualizing the otherwise intricate expansion pattern of this  planetary nebula. The global morphology of NGC 6210 is  reminiscent of other planetary nebulae with extended X-ray emission, suggesting the presence of a central hot bubble produced by shocked stellar wind. In spite of the dramatic structure of this PN substantial expanding radial motions are only found in the material surrounding the central, inner shell. The relatively slow radial motions of the collimated outflows and the bullet-like knots indicate that they are  moving away from the core very close to the plane of the sky, nearly perpendicular to the line of sight. The direction and ages of the various components, derived from their expanding velocities indicate that NGC 6210 has been strongly influenced by the  action of a bipolar, rotating, episodic outflow.
\end{abstract}

@@

\section{INTRODUCTION}
\label{sec:introduction}
NGC 6210 is a bright and relatively large PN in the northern sky. The first photographic image of NGC 6210 was published by \citet{Duncan:1937a}. In that old, fuzzy image a pair of twisted arms protruding in opposite directions from a blurred,  bright, nebular core are apparent. Several studies have discussed the inner kinematic motions of this object, e.g. \citet{Osterbrock:1966a, Weedman:1968a, Becker:1984a, Icke:1989a}. However, none of those studies had the spatial coverage and spectral resolution needed to produce a reliable spatio-kinematic model of this PN. 
The best kinematic study to date on NGC 6210 is probably the one published by \citet{Phillips:1996a}  who obtained 11 long-slits centered on the nebular core, with each slit at different position angles thus providing a good spatial coverage but with only limited, low and medium spectral resolution. From these data they suggested a reasonable, though somewhat basic, kinematic model for the ionized nebula. NGC 6210 jumped into fame thanks to the {\it HST} images obtained during the 1996 - 1998 period under the observing programs 6347 (K. Borkowski), 6792 (R. Rubin), 7501 (A. Hajian)  and 11122 (B. Balick). These images showed for the first time the extraordinarily complex morphology of NGC 6210. Incidentally the  {\it HST} and ground-based imagery prompted the nickname "the turttle" for this nebula, given the shape of its main body and extended arms that resemble the shell of a turttle with a long neck and its fins. The distance estimates to NGC 6210 vary roughly from 1.5 to 2.0 kpc. \citep{Hajian:1995a} derive a distance to NGC 6210 from VLA 6 cm  expansion parallax measurements of 1.57$\pm0.40$ kpc, considering an expansion velocity of 23 \kms, whereas Frew (20XX) derives a distance of kpc to this object from its H$\alpha$ surface brightness method.

@@

In Section 2, we describe our observations and the data-reduction steps, In Section 3 we describe the proper motions, In Section 4, we discuss the kinematic components from slit spectra. In next section, we present the dynamical ages and finally in section 5 we present the discussion and conclusion.

@@

Long-slit, echelle, spectroscopic observations of the nebula NGC~6210
were performed with the 2.1~m telescope at the Observatorio
Astron\'omico Nacional at San Pedro M\'artir, (OAN-SPM), Baja
California, M\'exico. We used the Manchester Echelle Spectrometer
(MES-SPM) \citep{Meaburn:2003a} on the 2.1 m telescope in its $f$/7.5
configuration.  The MES-SPM is a long-slit, echelle spectrometer that
has no cross-disperser; it isolates single orders using interference
filters. We used a 90\,\AA\, and 50\,\AA\, bandwidth
filter to isolate the 87th and 114th orders containing the
\Ha\,$+$\,\NIIlam\, and \OIIIlam\, nebular emission lines,
respectively. Spectroscopic data on NGC 6210 was collected within our regular observing programs over nine observing runs from 1998 till 2019. Table 1 shows the log of observations. In the first column the dates for the individual observing runs are listed; the second column lists the exposure times mostly used in the corresponding run; the third column lists the spectral range of the observations, namely, $H\alpha$ + [N II] and/or [O III]; the fourth column lists the slit width in microns and the fifth column the position angle of the slit. Finally, in the sixth column the nomenclature used to identify the slits obtained in each run is indicated. In the end 34 slit positions from all the runs were selected to analyse the kinematic structure of NGC 6210. The location of these slits are presented and labeled against an [N II] $\rm HST$ archive image in Figure 1. In the top panel the vertical solid lines represent the set of 11 slits obtained in the 2015 run. The vertical dashed lines represent 12 slit positions from other runs, in some cases they are highly coincident in location with the 2015 set but selected for reasons such as different exposure times or spectral range. The middle panel shows 5 slit positions oriented E - W and the bottom panel 6 slit positions obtained at an angle intended to record specific features, such as collimated outflows or the twisted arms of the nebula. In order to establish the exact position of the slit in each pointing, the slit position on the sky was recorded with an automatic procedure available in MES-SPM prior to the spectroscopic exposure.

During the 11 years span of the observations different detectors were used with MES. Their general characteristics are described below

A TEK-1 CCD detector was used for the 1998 data set with 1024 $\times$ 1024 square pixels, each 24$\mu$m on a side, $\equiv$0.312 arcsec pixel$^{-1}$. A 2$\times$2 binning was set, consequently, 512 increments, each 0\farcs624{} long gave a projected slit length of 5\farcm32 on the sky. For slits 150 $\mu$m ($\equiv$1.9\arcsec) and 70 $\mu$m ($\equiv$0.95\arcsec) wide the spectral resolution is 11.5 \kms and 4.6 \kms, respectively).

For the 2003, 2004 and 2011 runs of the present observations, we used a SITE-3 CCD detector. That detector had the same dimensions (1024 $\times$ 1024 square pixels, each 24$\mu$m on a side) as the Tek-1.

For the 2011 and 2013 runs we used a Marconi CCD detector with 2048 $\times$ 2048 square pixels, each 13.5
$\mu$m on a side. The detector was set to a binning of 2 $\times$ 2 in both the spatial and spectra directions. Consequently, 1024
increments, each 0\farcs352{} long gave a projected slit length of 5\farcm47 on the sky. For a 70~$\mu$m{} a velocity resolution of 5.9 \kms{} is obtained in this case, whereas the 150~$\mu$m slit yields 11.9 \kms{}.

@@

Figure 3 shows the location of the slit in the halo of NGC 6210  against a deep ground-based image. The bottom panel shows the large scale view with the location of the slit indicated. The top right panel is an enhanced contrast image of the region that intersects the slit, the regions shows a knotty, curved structure. The top left panel shows the corresponding spectrum from the 2019 run.

Most of the spectra presented here are included in the San Pedro M\'artir Kinematic Catalogue of Galactic Planetary Nebulae and are publicly available as pdf or fits files (http://kincatpn.astrosen.unam.mx)
\citep{Lopez:2012a}.

@@

\begin{table}
\centering
\caption{OAN-SPM/Mezcal high-resolution spectroscopy of NGC~6210}
\label{table:pa5}
\begin{tabular}{l|cccccccc} \toprule
  &   Exp.~Time & Spectral  & Slit & P. A.   & slit        \\
  Epoch   &    (s) & Range  &  (\AA)    &    (\AA)  & (m$\mu$)     \\
  \midrule
1998/06/28          & 30 & \Ha  & 150 & 90 & O\\
1998/06/28          & 300 & \Ha  & 150 & 90 & N\\
1998/06/28           & 1200 & \Ha   & 150 & 90 & M,P,Q \\
2003/06/05         & 1800 &  \Ha   & 70 &0 &     A,B,D*,E*,F,H,I,K\\ 
2003/10/16   &  1800 &   \Ha    & 70 & $-$21 & T\\
2003/10/16    &  1800 &   \Ha   & 70 & $-$68 & V\\
2003/10/17    &  1800 &   \Ha  & 70 & 77 & W\\
2004/06/13   & 1800 &   \oiii  & 70 & $-$9 & R  \\  
2004/06/13   & 1800 &   \oiii  & 70 & $-$19 & S  \\  
2004/06/14    & 1800 &  \sii &150 & $-$19 & S  \\
2004/06/13   & 1800 &   \oiii  & 70 & $-$56 & U  \\  
2004/06/14   & 1800 &   \oiii  & 150 & 0 & A',L  \\
2011/05/21     & 1800 & \Ha  & 150 & 0 & G* \\
2011/05/21     & 600 & \oiii & 150 & 0  & G \\
2013/07/06    &  1800 & \Ha   & 150 & 0 & C,I,J \\
2015/08/18  &  1800 &   \Ha, \oiii  & 70 & 0 & C,D,E,F,G\\
2015/08/19  &  1800 &   \Ha, \oiii   & 70 & 0 & B,A*,I\\
2015/08/20   &  1800 &   \Ha, \oiii & 70 & 0 & H,J,K\\
2019/09/18  &  1800 &   \Ha    & 150 & 56 & \\
  \bottomrule
\end{tabular}
\end{table}

@@

Using FORTRAN and phython routines, we produced  velocity cubes in  \NIIlam{} and \oiii{} emission lines, in order to 
...

The resulting data cubes for the \Halam\, and \NIIlam\, lines are shown
in Figure 3 as isovelocity channel maps, each isovelocity map is 20
\kms\, wide.

@@

The miscellaneous high-ionization components lie outside the inner shell
and show a spoke-like morphology on the \textit{HST} images.
They are represented by dot-dashed ellipses in Figure~\ref{fig:shell-velocity-components} and small green triangle symbols in Figure~\ref{fig:shell-velocity-axes}.
They do not show any marked kinematic pattern,
but are broadly compatible with the velocities of nearby portions of the intermediate shell.

@@

The \nii{} emission from the nebula is dominated by small-scale knots,
as can be appreciated on the \textit{HST} images (Figure~\ref{fig:hst}).
These knots have typical sizes \(< 1''\) and so are not spatially resolved
in our ground-based spectroscopy.
A small number of isolated knots are individually detected in our spectra,
but in general we detect only the combined emission of extended knot complexes.
Figure~\ref{fig:knot-complex-map} shows all the \nii{} emission components
identified from our spectra, but excluding those associated with the high-ionization
shells discussed above.
As previously, the components are divided into negative (left panel)
and positive (right panel) velocities.  
In Figure~\ref{fig:knot-complex-velocity-axes}
we classify the emission components into five knot complexes,
plus three individual knots, whose plane-of-sky distribution
and typical line-of-sight velocity are shown
superimposed on the \textit{HST} \nii{} image (contours)
and isovelocity channel maps reconstructed from our slit spectra (color images).

Figure~\ref{fig:knot-complexes} shows the velocity of each \nii{} component
as a function of position along the two axes that characterize the high-ionization shells:
axis~A at \(\text{PA} = \ang{160}\)
and axis~B at \(\text{PA} = \ang{70}\) (see \S~\ref{sec:high-ioniz-shells}).
The kinematics of the shells themselves are shown by faint continuous lines for comparison.
Note that the extent of the velocity axis in this figure
is twice as large as in Figure~\ref{fig:shell-velocity-axes}.

The clearest axial alignment is seen for the SW Red complex,
which is very tightly clustered around the negative arm of axis~B.
The same is seen to a lesser extent for the NE Blue complex
around the positive arm of axis~B,
but the distribution around the axis is broader in this case.
The remaining complexes have no clear alignment with either axis,
except that the SE Blue complex is marginally more elongated
along the positive arm of axis~A.

@@

To further investigate the structure of NGC~6210,
we have used HST WFPC2 narrow-band images to derive the surface brightness in the lines \Ha{} \Wav{6563}, \Hb{} \Wav{4861}, \nii{} \Wav{6583}, and \oiii{} \Wav{5007}.
Flux calibration is carried out following the procedure outlined in \citet{Rubin:2002a}, which accounts for contamination of each emission line filter by non-target lines and atomic continuum emission \citetext{see also \citealp{Ueta:2019a}}.
Foreground dust extinction is corrected for using the observed Balmer decrement, \(\Ha/\Hb\),
assuming an intrinsic value of 2.85 and the reddening law of \citet{Cardelli:1989a},
yielding an average extinction of \(C(\Hb) = 0.13\) for the nebula.
The total intrinsic \Ha{} flux for the nebula is then found by summing over the entire HST image,
yielding \(F(\Ha) = \SI{3.176e-10}{erg.s^{-1}.cm^{-2}}\),
which is almost identical to that found by previous studies \citep{Liu:2004a}.

@@

What is the origin of the asymmetries?
All mechanisms proposed to date \citet{Soker:2004b, Soker:2016b}
involve accretion disk and jet formation in a triple star system.
This may be via accretion of the AGB wind onto a close binary on a wide orbit around the AGB,
or, if the orbit is closer, via a common envelope or grazing envelope phase,
which may or may not result in the merger of the close binary pair.
The descendant system would be binary if a merger or ejection occurred,
or still a triple if not,
but there is no direct evidence that the Turtle's central star is binary or triple.
The fact that the basic parameters of the nebula and its central star can be consistently interpreted in terms of single-star evolution (see \S~\ref{sec:comparison-with-post})
suggests that a triple interaction was not required for the initial ejection of the envelope that forms the intermediate shell,
although it may be responsible for its subsequent shaping.

@@

It is particularly instructive to compare the Turtle with the almost-spherical NGC~1535,
since the central stars of these two nebulae are almost perfect twins \citep{Herald:2011a},
with closely similar effective temperature, surface gravity, and stellar wind mass-loss rate and velocity.
The basic parameters of the nebulae
(radii, densities, chemical composition, and velocities of the shells and halo)
are also very similar.\footnote{
  Minor exceptions are that NGC~1535 is slightly higher excitation,
  as measured by \heii{}/\Hb{} \citep{Barker:1989a},
  and has a higher carbon abundance \citep{Kwitter:1996a},
  although both nebulae are oxygen-rich with \chem{C/O < 1}.
}
Despite these similarities,
NGC~1535 is totally lacking the fast lobes,
low-ionization knots, and asymmetries that are so prominent in the Turtle.
The implication is that none of these phenomena are intrinsic or necessary
for the formation and expansion of the planetary nebula.

>>>>

\begin{abstract}
  We carry out a comprehensive kinematic and morphological study
  of the asymmetrical planetary nebula:
  NGC~6210, known as the Turtle.
  The nebula's spectacularly chaotic appearance has led to proposals that it was shaped by mass transfer in a triple star system.
  We study the three-dimensional structure and kinematics
  of its shells, lobes, knots, and haloes
  by combining radial velocity mapping from multiple long-slit spectra
  with proper motion measurements from multi-epoch imaging.
  We find that the nebula has five distinct ejection axes.
  The first is the axis of the bipolar, wind-blown inner shell,
  while the second is the axis of the lop-sided, elliptical, fainter, but more massive intermediate shell.
  A further two axes are bipolar flows that form the point symmetric,
  high-ionization outer lobes, all with inclinations close to the plane of the sky.
  The final axis, which is inclined close to the line of sight,
  traces collimated outflows of low-ionization knots.
  We detect major changes in outflow directions over the last 6000~years,
  beginning in the pre-planetary nebula phase
  and continuing through the ionization of the nebula, which began 3500~years ago.
  Most notably,
  all redshifted low-ionization knots have kinematic ages greater than 2000~years,
  whereas all blueshifted knots have ages younger than 2000~years. 
  Such a sudden and permanent 180-degree flip in the ejection axis
  at a relatively late stage in the nebular evolution is a challenge to models of planetary nebula formation and shaping.
  % A comprehensive kinematic study has been carried out on the planetary nebula (PN) NGC 6210. Multiple long-slits, echelle spectra have been obtained over the face of this nebula mapping its full shell structure, the opposite pairs of extended, twisted arms and the bipolar collimated outflows and bullets. Public {\it HST} imagery has been used to identify kinematic elements with structural components. The long-slit spectroscopic information has been combined into channel maps that greatly facilitate visualizing the otherwise intricate expansion pattern of this  planetary nebula. The global morphology of NGC 6210 is  reminiscent of other planetary nebulae with extended X-ray emission, suggesting the presence of a central hot bubble produced by shocked stellar wind. In spite of the dramatic structure of this PN substantial expanding radial motions are only found in the material surrounding the central, inner shell. The relatively slow radial motions of the collimated outflows and the bullet-like knots indicate that they are  moving away from the core very close to the plane of the sky, nearly perpendicular to the line of sight. The direction and ages of the various components, derived from their expanding velocities indicate that NGC 6210 has been strongly influenced by the  action of a bipolar, rotating, episodic outflow.
\end{abstract}

@@

\section{INTRODUCTION}
\label{sec:introduction}
NGC 6210 is a bright and relatively large PN in the northern sky. The first photographic image of NGC 6210 was published by \citet{Duncan:1937a}. In that old, fuzzy image a pair of twisted arms protruding in opposite directions from a blurred,  bright, nebular core are apparent. Several studies have discussed the inner kinematic motions of this object, e.g. \citet{Osterbrock:1966a, Weedman:1968a, Becker:1984a, Icke:1989a}. However, none of those studies had the spatial coverage and spectral resolution needed to produce a reliable spatio-kinematic model of this PN. 
The best kinematic study to date on NGC 6210 is probably the one published by \citet{Phillips:1996a}  who obtained 11 long-slits centered on the nebular core, with each slit at different position angles thus providing a good spatial coverage but with only limited, low and medium spectral resolution. From these data they suggested a reasonable, though somewhat basic, kinematic model for the ionized nebula. NGC 6210 jumped into fame thanks to the {\it HST} images obtained during the 1996 - 1998 period under the observing programs 6347 (K. Borkowski), 6792 (R. Rubin), 7501 (A. Hajian)  and 11122 (B. Balick). These images showed for the first time the extraordinarily complex morphology of NGC 6210. Incidentally the  {\it HST} and ground-based imagery prompted the nickname "the turttle" for this nebula, given the shape of its main body and extended arms that resemble the shell of a turttle with a long neck and its fins. The distance estimates to NGC 6210 vary roughly from 1.5 to 2.0 kpc. \citep{Hajian:1995a} derive a distance to NGC 6210 from VLA 6 cm  expansion parallax measurements of 1.57$\pm0.40$ kpc, considering an expansion velocity of 23 \kms, whereas \citet{Frew:2016a} derives a distance of \SI{2.1 \pm 0.5}{kpc} to this object from its H$\alpha$ surface brightness method.

@@

In \S~\ref{sec:observations}, we describe our spectroscopic observations and corresponding data-reduction steps.
In \S~\ref{sec:proper-motions} we describe the proper motions that we derive from archival \textit{HST} images.
In \S~\ref{sec:kinematic-components} we extract kinematic components from the slit spectra and classify them into different systems under the categories of shells, knots, lobes and halo, which are each analysed.
In \S~\ref{sec:three-dimens-struct} we combine the proper motion and spectroscopic observations to reconstruct the three-dimensional structure of the nebula and derive its mass-loss history.
In \S~\ref{sec:discussion} we discuss our results in the broader context of planetary nebula studies and post-AGB stellar evolution.
Finally, in \S~\ref{sec:conclusions} we summarise our conclusions.
Technical details of the per-slit velocity components,
density calibration,
and ionization diagnostics are given in a series of appendices.

@@

Long-slit, echelle, spectroscopic observations of the nebula NGC~6210
were performed with the 2.1~m telescope at the Observatorio
Astron\'omico Nacional at San Pedro M\'artir, (OAN-SPM), Baja
California, M\'exico. We used the Manchester Echelle Spectrometer
(MES-SPM) \citep{Meaburn:2003a} on the 2.1 m telescope in its $f$/7.5
configuration.  The MES-SPM is a long-slit, echelle spectrometer that
has no cross-disperser; it isolates single orders using interference
filters. We used a 90\,\AA\, and 50\,\AA\, bandwidth
filter to isolate the 87th and 114th orders containing the
\Ha\,$+$\,\NIIlam\, and \OIIIlam\, nebular emission lines,
respectively. Spectroscopic data on NGC 6210 was collected within our regular observing programs over nine observing runs from 1998 till 2019. Table 1 shows the log of observations. In the first column the dates for the individual observing runs are listed; the second column lists the exposure times mostly used in the corresponding run; the third column lists the spectral range of the observations, namely, $H\alpha$ + [N II] and/or [O III]; the fourth column lists the slit width in microns and the fifth column the position angle of the slit. Finally, in the sixth column the nomenclature used to identify the slits obtained in each run is indicated. In the end 34 slit positions from all the runs were selected to analyse the kinematic structure of NGC 6210. The location of these slits are presented and labeled against an \oiii{} \textit{HST} archive image in Figure~\ref{fig:slit-positions}. In the top panel the vertical solid lines represent the set of 11 slits obtained in the 2015 run. The vertical dashed lines represent 12 slit positions from other runs, in some cases they are highly coincident in location with the 2015 set but selected for reasons such as different exposure times or spectral range. The middle panel shows 5 slit positions oriented E - W and the bottom panel 6 slit positions obtained at an angle intended to record specific features, such as collimated outflows or the twisted arms of the nebula. In order to establish the exact position of the slit in each pointing, the slit position on the sky was recorded with an automatic procedure available in MES-SPM prior to the spectroscopic exposure.

During the 11 year span of the observations, we used different detectors with MES, 
with general characteristics as follows.
A \mbox{TEK-1} CCD detector was used for the 1998 observations,
consisting of \(1024 \times 1024\) square pixels, 
each \SI{24}{\um} on a side,
used with \(2\times2\) binning to give a plate scale of \SI{0.624}{arcsec} per double-binned pixel
and a projected slit length of \SI{5.32}{arcmin}.
The spectral resolution was \SI{11.5}{km.s^{-1}} for the \SI{150}{\um} (\SI{1.9}{arcsec}) slit
and \SI{4.6}{km.s^{-1}} for the \SI{70}{\um} (\SI{0.95}{arcsec}) slit.
For the 2003, 2004, and 2011 observations, we used a \mbox{SITE-3} CCD detector with the same dimensions as the \mbox{TEK-1}. 
For the 2011 and 2013 runs we used a Marconi CCD detector
with \(2048 \times 2048\) square pixels, each \SI{13.5}{\um} on a side.
Again, \(2\times2\) binning was used to give a plate scale of \SI{0.352}{arcsec} per double-binned pixel
and a projected slit length of \SI{5.47}{arcmin},
with a spectral resolution of \SI{11.9}{km.s^{-1}} (\SI{150}{\um} slit)
or \SI{5.9}{km.s^{-1}} (\SI{70}{\um} slit).

@@

Figure 3 shows the location of the slit in the halo of NGC 6210  against a deep ground-based image. The bottom panel shows the large scale view with the location of the slit indicated. The top right panel is an enhanced contrast image of the region that intersects the slit, the regions present a knotty, curved structure. The top left panel shows the corresponding spectrum from the 2019 run.

Most of the spectra presented here are included in the \mbox{San Pedro Mártir} Kinematic Catalogue of Galactic Planetary Nebulae\footnote{
  \url{http://kincatpn.astrosen.unam.mx}
}
and are publicly available as pdf or fits files
\citep{Lopez:2012a}.

@@

\begin{table}
\centering
\caption{OAN-SPM/Mezcal high-resolution spectroscopy of NGC~6210}
\label{table:pa5}
\begin{tabular}{@{}l@{} rrrr l@{}} \toprule
  &   Exp.~Time & Spectral  & Slit & P.A.   & \\
  Epoch   &    (s) & Range  &  (\AA)    &    (\(^\circ\))  & Slit Label \\
  \midrule
1998/06/28 & 30 & \Ha  & 150 & 90 & O\\
1998/06/28 & 300 & \Ha  & 150 & 90 & N\\
1998/06/28 & 1200 & \Ha   & 150 & 90 & M,P,Q \\
2003/06/05 & 1800 &  \Ha   & 70 &0 &     A,B,D*,E*,F,H,I,K\\ 
2003/10/16 &  1800 &   \Ha    & 70 & $-$21 & T\\
2003/10/16 &  1800 &   \Ha   & 70 & $-$68 & V\\
2003/10/17 &  1800 &   \Ha  & 70 & 77 & W\\
2004/06/13 & 1800 &   \oiii  & 70 & $-$9 & R  \\  
2004/06/13 & 1800 &   \oiii  & 70 & $-$19 & S  \\  
2004/06/14 & 1800 &  \sii &150 & $-$19 & S  \\
2004/06/13 & 1800 &   \oiii  & 70 & $-$56 & U  \\  
2004/06/14 & 1800 &   \oiii  & 150 & 0 & A',L  \\
2011/05/21 & 1800 & \Ha  & 150 & 0 & G* \\
2011/05/21 & 600 & \oiii & 150 & 0  & G \\
2013/07/06 &  1800 & \Ha   & 150 & 0 & C,I,J \\
2015/08/18 &  1800 &   \Ha, \oiii  & 70 & 0 & C,D,E,F,G\\
2015/08/19 &  1800 &   \Ha, \oiii   & 70 & 0 & B,A*,I\\
2015/08/20 &  1800 &   \Ha, \oiii & 70 & 0 & H,J,K\\
2019/09/18 &  1800 &   \Ha    & 150 & 56 & X\\
  \bottomrule
\end{tabular}
\end{table}

@@

% Using FORTRAN and phython routines, we produced  velocity cubes in  \NIIlam{} and \oiii{} emission lines, in order to 
% ...

% The resulting data cubes for the \Halam\, and \NIIlam\, lines are shown
% in Figure 3 as isovelocity channel maps, each isovelocity map is 20
% \kms\, wide.

@@

The miscellaneous high-ionization components lie outside the inner shell,
% and show a spoke-like morphology on the \textit{HST} images.
and are represented by dot-dashed ellipses in Figure~\ref{fig:shell-velocity-components}
and small green triangle symbols in Figure~\ref{fig:shell-velocity-axes}.
They do not show any marked kinematic pattern,
but are broadly compatible with the velocities of nearby portions of the intermediate shell.

@@

The \nii{} emission from the nebula is dominated by small-scale knots,
as can be appreciated on the \textit{HST} images (Figure~\ref{fig:hst}).
These knots have typical sizes \(< 1''\) and so are not spatially resolved
in our ground-based spectroscopy.
A small number of isolated knots are individually detected in our spectra,
but in general we detect only the combined emission of extended knot complexes.
Figure~\ref{fig:knot-complex-map} shows all the \nii{} emission components
identified from our spectra, but excluding those associated with the high-ionization
shells discussed above.
As previously, the components are divided into negative (left panel)
and positive (right panel) velocities.  
In Figure~\ref{fig:knot-complexes}
we classify the emission components into five knot complexes,
plus three individual knots, whose plane-of-sky distribution
and typical line-of-sight velocity are shown
superimposed on the \textit{HST} \nii{} image (contours)
and isovelocity channel maps reconstructed from our slit spectra (color images).

Figure~\ref{fig:knot-complex-velocity-axes} shows the velocity of each \nii{} component
as a function of position along the two axes that characterize the high-ionization shells:
axis~A at \(\text{PA} = \ang{160}\)
and axis~B at \(\text{PA} = \ang{70}\) (see \S~\ref{sec:high-ioniz-shells}).
The kinematics of the shells themselves are shown by faint continuous lines for comparison.
Note that the extent of the velocity axis in this figure
is twice as large as in Figure~\ref{fig:shell-velocity-axes}.

% The clearest axial alignment is seen for the SW Red complex,
% which is very tightly clustered around the negative arm of axis~B.
% The same is seen to a lesser extent for the NE Blue complex
% around the positive arm of axis~B,
% but the distribution around the axis is broader in this case.
% The remaining complexes have no clear alignment with either axis,
% except that the SE Blue complex is marginally more elongated
% along the positive arm of axis~A.

@@

To further investigate the structure of NGC~6210,
we have used HST WFPC2 narrow-band images to derive the surface brightness in the lines \Ha{} \Wav{6563}, \Hb{} \Wav{4861}, \nii{} \Wav{6583}, and \oiii{} \Wav{5007}.
Flux calibration is carried out following the procedure outlined in \citet{Rubin:2002a}, which accounts for contamination of each emission line filter by non-target lines and atomic continuum emission \citetext{see also \citealp{Ueta:2019a}}.
Foreground dust extinction is corrected for using the observed Balmer decrement, \(\Ha/\Hb\),
assuming an intrinsic value of 2.85 and the reddening law of \citet{Cardelli:1989a},
yielding an average extinction of \(C(\Hb) = 0.13\) for the nebula.
The total extinction-corrected \Ha{} flux for the nebula is then found by summing over the entire HST image,
yielding \(F(\Ha) = \SI{3.176e-10}{erg.s^{-1}.cm^{-2}}\),
which is very close to that found by previous studies \citep{Liu:2004a}.

@@

What is the origin of the asymmetries?
Mechanisms proposed to date \citet{Soker:2004b, Soker:2016b}
involve accretion disk and jet formation in a triple star system.
This may be via accretion of the AGB wind onto a close binary on a wide orbit around the AGB,
or, if the orbit is closer, via a common envelope or grazing envelope phase,
which may or may not result in the merger of the close binary pair.
The descendant system would be binary if a merger or ejection occurred,
or still a triple if not,
but there is no direct evidence that the Turtle's central star is binary or triple.
The fact that the basic parameters of the nebula and its central star can be consistently interpreted in terms of single-star evolution (see \S~\ref{sec:comparison-with-post})
suggests that a triple interaction was not required for the initial ejection of the envelope that forms the intermediate shell,
although it may be responsible for its subsequent shaping.

@@

It is particularly instructive to compare the Turtle with the almost-spherical NGC~1535,
since the central stars of these two nebulae are almost perfect twins \citep{Herald:2011a},
with closely similar effective temperature, surface gravity, and stellar wind mass-loss rate and velocity.
The basic parameters of the nebulae
(radii, densities, chemical composition, and velocities of the shells and halo)
are also similar.\footnote{
  Minor exceptions are that NGC~1535 is slightly higher excitation,
  as measured by \heii{}/\Hb{} \citep{Barker:1989a},
  and has a higher carbon abundance \citep{Kwitter:1996a},
  although both nebulae are oxygen-rich with \chem{C/O < 1}.
}
Despite these similarities,
NGC~1535 is totally lacking the fast lobes,
low-ionization knots, and asymmetries that are so prominent in the Turtle.
The implication is that none of these phenomena are intrinsic or necessary
for the formation and expansion of the main planetary nebula shell.
